---
title: "00. Circular Bar Plot"
author: "Babatunde Adebayo"
date: "8/21/2020"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recreation Attempt
```{r setup}
library(tidyverse)
library(extrafont)
library(ggthemes)
```

```{r data}
ng_today <- readxl::read_excel("../../data/Count/Daily Count By State_final_V2.xlsx", sheet = "Today")
```

### Explore Data
Attempt to use {skimr} package to get a sense of overview about the size of your data, missing values, mean rates, and a basic histogram to show the distribution of numerical data-type columns.
```{r exploration}
skimr::skim(ng_today)
```
```{r}
rsummary <- ng_today %>% 
  group_by(Geo_Zone) %>% 
  summarise(z_tot = sum(Tot_Cases)) %>% 
  ungroup() %>% 
  arrange(desc(z_tot)) 

19815/sum(ng_today$Tot_Cases)
```


### Data Processing

```{r data_prep}
# data_scaled 
ng_today_uscal <- ng_today %>% 
  # left_join(rsummary) %>% 
  mutate(Geo_Zone = factor(Geo_Zone,levels = c("South West", "North Central", "South South", "North West", "South East", "North East"))) 

ng_today_scal <- ng_today %>% 
  mutate(Geo_Zone = factor(Geo_Zone,levels = c("South West", "North Central", "South South", "North West", "South East", "North East")),
         Tot_Cases = scales::rescale(Tot_Cases, to = c(0, 100)))

```

### Plotting Setup
The goal is to create circular barplot plot with extra-sauce of grouped format structure. The extra-sauce here include some extra row space to separate groups, arc line for group labelling and smooth grid for y-axis labelling. 
Hence, there is a need for some data preps and geometry calculations (not to worry on this data-to-viz to the rescue).

The basic steps are as follow:
  1. Create empty bar rows in the data to allow grouping 
2. Build labels and add to data. In this case, we will label the State per bar and use continent for the grouping. More importantly for angular geometry for labels rotation.
3. Base Data - Dataframe handles continent labelling and line arc aesthetics. 
4. Grid Data - A dataframe to structure y-axis scale (It basically acts as the y-axis label) -  in the case of the scaled data it will handle bar of 0-20-40-60-80 

### Unscaled Data
```{r}
empty_bar=4
to_add = data.frame(matrix(NA, empty_bar*nlevels(ng_today_uscal$Geo_Zone), ncol(ng_today_uscal)) )

colnames(to_add) = colnames(ng_today_uscal)
to_add$Geo_Zone = rep(levels(ng_today_uscal$Geo_Zone), each=empty_bar)

data_uscl <- full_join(ng_today_uscal, #%>% arrange(desc(z_tot), Geo_Zone), 
                        to_add)   %>%  #rbind(data, to_add)
  left_join(rsummary) %>% 
  arrange(desc(z_tot)) %>%
  mutate(id = row_number())

# Step 2
# Get the name and the y position of each label
label_data=data_uscl
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
label_data$Tot_Cases2 = ifelse(label_data$Tot_Cases < max(label_data$Tot_Cases,na.rm = T), label_data$Tot_Cases+500,label_data$Tot_Cases)


# step 3
# prepare a data frame for base lines
base_data2 = data_uscl %>% 
  group_by(Geo_Zone) %>% 
  # arrange(desc(Geo_Zone), desc(z_tot) ) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)),
         g_label = gsub(" ","\n",as.character(Geo_Zone)))

# Step 4
# prepare a data frame for grid (scales)
grid_data2 <- tibble(start = c(7,18,28,39,48) +0.5,
                     end = c(9,20,30,41,50)-0.5,
                     title = mean(c(start,end)))
grid_data2
```

```{r build_plot, fig.height=8, fig.width=7}
d_col = "white"
alp = 1

cirl_plt <- 
  data_uscl %>% 
ggplot(aes(x=as.factor(id), y=Tot_Cases, fill=Geo_Zone)) + #NB id variable is a factor
  geom_bar(stat = "identity", alpha =0.5) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sure barplots are OVER it.
  geom_segment(data=grid_data2, aes(x = end, y = 20000, xend = start, yend = 20000), colour = d_col, alpha=alp, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data2, aes(x = end, y = 15000, xend = start, yend = 15000), colour = d_col, alpha=alp, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data2, aes(x = end, y = 10000, xend = start, yend = 10000), colour = d_col, alpha=alp, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data2, aes(x = end, y = 5000, xend = start, yend = 5000), colour = d_col, alpha=alp, size=0.3 , inherit.aes = FALSE ) +
  # Add text showing the value of each 100/75/50/25 lines # Modify this accordingly for an unscaled data
  annotate("text", x = rep(max(data_uscl$id),4), y = c(5000, 10000, 15000, 20000), label = c("5,000", "10,000", "15,000", "20,000") , color=d_col, size=3 , angle=0, fontface="bold", hjust=1) +
  # geom_bar(aes(x=as.factor(id), y=Tot_Cases, fill=Geo_Zone), stat="identity", alpha=0.5) +
  ylim(-20000,20000) + 
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(), # Modify this to keep x or y axis text
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "black", colour = NA), #element_rect(fill = "black"),
    plot.title = element_text(hjust = 0.5, size = 24, family  = "Garamond"),
    plot.subtitle = element_text(hjust = 0.5, family  = "Garamond"),
    plot.caption = element_text(hjust = 0.5, family = "Garamond")
    # plot.margin = unit(rep(-1,4), "cm") <- this hides the plot title if not commented out
  ) + 
  coord_polar() + # Make it a circle
  # add labels - save to remove inherit.aes = FALSE (soft-deprecated I believe)
  geom_text(data=label_data, aes(x=id, y=Tot_Cases2, label=State, hjust=hjust), color=d_col, fontface="bold",alpha=0.7, size=3.5, angle= label_data$angle) #+
  # Add base line information - Arc (a line in a polar coordinate) Then Label ( gridText Can be use to enhance text placement - say for a two line display)
  # geom_segment(data=base_data2, aes(x = start, y = -450, xend = end, yend = -450), colour = "white", alpha=0.8, size=0.7 , inherit.aes = FALSE )  #+
  # geom_text(data=base_data2, aes(x = title, y = -18, label=g_label),
  # hjust=c(1.2,0,0,0,0,0), colour = "white", alpha=0.8, size=4,
  # fontface="bold", inherit.aes = FALSE)


ggsave(plot = cirl_plt, filename = "./Poster/png/nig_circular_plot_v2.png", device = 'png',dpi = 320)
```

